A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array

### 2015-01-03 ms.tex
Minor changes to notations, citations, and wording. Added a file addressing referee comments

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.  Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space. Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.

@@

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,bea13,dil13,thy13,pob14}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13} and \citet{par14}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify this contamination by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique. In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_{bf}(\vec{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\vec{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{s},f)$ and $A(\hat{s},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{s}$), and $W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}(\vec{b},f)$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
\begin{align}\label{eqn:obsvis}
  V_{bf}(\vec{b},f) &= \iint\limits_\textrm{sky} A(\hat{s},f)\,I(\hat{s},f)\,W_f(f)\,e^{-i2\pi f\frac{\vec{b}\cdot\hat{s}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element along $\hat{s}$. We wish to emphasize that this equation is valid in general and does not involve any approximation.

We define the {\it delay spectrum}, $V_{b\tau}(\vec{b},\tau)$, to be the inverse Fourier transform of $V_{bf}(\vec{b},f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_{b\tau}(\vec{b},\tau) &= \int V_{bf}(\vec{b},f)\,W'_f(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\vec{b}\cdot\hat{s}}{c}.
\end{equation}

Equation~\ref{eqn:obsvis} can be modified to express the Fourier decomposition of transverse sky brightness distribution as:
\begin{align}\label{eqn:vis}
  V_{uf}(\vec{u},f) &= \iint\limits_\textrm{sky} A(\hat{s},f)\,I(\hat{s},f)\,W_f(f)\,e^{-i2\pi\vec{u}\cdot\hat{s}}\,\dif\Omega,
\end{align}
where, $\vec{u}$ denotes the transverse spatial frequency vector. $\vec{u}$ is directly related to $\vec{b}$ as $\vec{u}\equiv \vec{b}/\lambda$, and to the transverse spatial wavenumber mode, $\vec{k}_\perp$, as $\vec{k}_\perp\equiv 2\pi\vec{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$. Hence, $V_{uf}(\vec{u},f)\equiv V_{uf}(\vec{b}/\lambda,f)\equiv V_{bf}(\lambda\vec{u},f)$. It is important to note that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.

Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight. $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time. It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. This approximation holds under the assumption that there is negligible cosmological evolution within the redshift range (or frequency band). Thus,
\begin{align}\label{eqn:los-transform}
  V_{u\eta}(\vec{u},\eta) &= \int V_{uf}(\vec{u},f)\,W'_f(f)\,e^{i2\pi f\eta}\,\dif f
\end{align}
represents the true spatial Fourier representation of the three--dimensional sky brightness distribution. This representation has been discussed in detail in \citet{mor04}. The spatial power spectrum of EoR H{\sc i} distribution, $P(\vec{k}_\perp,k_\parallel)$, and $V_{u\eta}(\vec{u},\eta)$ are related by $P(\vec{k}_\perp,k_\parallel)\propto |V_{u\eta}(\vec{u},\eta)|^2$.

While $V_{b\tau}(\vec{b},\tau)$ and $V_{u\eta}(\vec{u},\eta)$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical. The difference basically arises from the fact noted above that $\vec{b}$ is independent of frequency while $\vec{u}$ is not. \citet{par12} and \citet{liu14a} have discussed the mathematical correspondence between the two.

@@

In summary, the delay spectrum, $V_{b\tau}(\vec{b},\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}(\vec{b},\tau)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power $|V_{u\eta}(\vec{u},\eta)|^2$ in Fourier modes containing critical information about spatial scales. For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}(\vec{b},\tau)|$.

Note that since {\it visibility} is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays. Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector. In this paper, $\vec{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively. Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\vec{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries are called horizon delay limits and are shown by solid lines. The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.

In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\vec{k}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function. We call this region with medium shade as the {\it EoR window} in this paper. The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\includegraphics[width=\linewidth]{f1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented. The $x$--axis is denoted by $|\vec{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the instrumental transfer function are expected to be relatively free of foreground contamination. Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\vec{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.

The MWA passband of width 30.72~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.

@@

We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect. In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s. We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, sorted by baseline length, $|\vec{b}|$. Note that visibilities from different baselines have not been averaged together in this analysis.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\includegraphics[width=\linewidth]{f3.eps}
\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:fhd_data}}
\end{figure}

@@

Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, simulated with only the diffuse foreground model. The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.

>>>>

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function. Using the delay spectrum technique {\bf that maps the Fourier transform of the measurements to locations of foreground objects on the celestial sphere through delays in signals received by antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space. Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.

@@

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. {\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}.

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}. % A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. 
The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it \textbf{Per-Baseline}} approach of \citet{par12b}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the {\bf shape of the delay spectrum} of each baseline. We report two important findings: the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify this contamination by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}. In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
\begin{align}\label{eqn:obsvis}
  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. We wish to {\bf note} that this equation is valid {\bf even for wide--field applications where the sky cannot be approximated to be flat}.

We define the {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, to be the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_{b\tau} &\equiv \int V_{bf}\,W'_f\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W'_f\equiv W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}

Equation~\ref{eqn:obsvis} can be {\bf equivalently re-expressed} as:
\begin{align}\label{eqn:vis}
  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
\end{align}
where, $\boldsymbol{u}$ denotes the transverse spatial frequency vector. $\boldsymbol{u}$ is directly related to $\boldsymbol{b}$ as $\boldsymbol{u}\equiv \boldsymbol{b}/\lambda$, and to the transverse spatial wavenumber mode, $\boldsymbol{k}_\perp$, as $\boldsymbol{k}_\perp\equiv 2\pi\boldsymbol{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$. Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$. It is important to note that for a given antenna pair, $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.

Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight. $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time. It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible. Thus,
\begin{align}\label{eqn:los-transform}
  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
\end{align}
represents the true spatial Fourier representation of the three--dimensional sky brightness distribution. This representation has been discussed in detail in \citet{mor04}. The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by $P(\boldsymbol{k}_\perp,k_\parallel)\propto |V_{u\eta}|^2$.

While $V_{b\tau}$ and $V_{u\eta}$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical. The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not. \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.

@@

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power $|V_{u\eta}|^2$ in Fourier modes containing critical information about spatial scales. For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}|$.

Note that since {\it visibility} is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays. Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector. In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively. Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries are called horizon delay limits and are shown by solid lines. The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.

In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function. We call this region with medium shade as the {\it EoR window} in this paper. The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\includegraphics[width=\linewidth]{f1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented. The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the instrumental transfer function are expected to be relatively free of foreground contamination. Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.

The MWA passband of width 30.72~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.

@@

We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect. In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s. We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12b}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$. Note that visibilities from different baselines have not been averaged together in this analysis.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\includegraphics[width=\linewidth]{f3.eps}
\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:fhd_data}}
\end{figure}

@@

Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}|$, simulated with only the diffuse foreground model. The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
